---
title: "Figure1"
author: "Riccardo Mosca"
date: "2025-05-12"
output:
  word_document: default
  html_document: default
---

This markdown show how to generate Figure 1B and Figure S1E-F

Figure 1B
Loading packages and qPCR fold changes values
```{r, message=FALSE, warning=FALSE,tidy=TRUE}
library(dplyr)
library(ggplot2)
library(ggpubr)
library(data.table)





```

loading library size for fig S1A
```{r}
read_length <- read.table(file = "/Users/riccardomosca/Downloads/read_lengths.txt", header = F, stringsAsFactors = F)
read_length$Name <- "Input"
colnames(read_length) <- c("length", "Name")
summary(read_length$length)
```
plotting
```{r}
p <- ggplot(read_length, aes(x=Name, y=length)) + 
  geom_violin(trim=FALSE, fill="grey", alpha=0.5) +
  geom_boxplot(width=0.1, fill="white", color="black", outlier.shape=NA) +
  coord_cartesian(ylim=c(0,150)) +
  theme_classic() +
  ylab("Input library insert size (nt)") +
  xlab("3.45x10**7 reads")

p
ggsave("/Users/riccardomosca/Desktop/RAPseq_PAPER/FIGUREs/FIGURE1/NEW/FigS1A.pdf", p, width = 3, height = 5)
```

Figure 1B
Loading qPCR fold changes values
```{r}
qPCR <- read.table(file = "/Users/riccardomosca/Desktop/RAPseq_PAPER/FIGUREs/FIGURE1/NEW/HUR_qPCR", header = T, stringsAsFactors = F)
qPCR_means <- qPCR %>% group_by(Substrate, Target) %>% summarise(Mean_Fold_Change = mean(Fold_Change))
qPCR_means <- as.data.frame(qPCR_means)
head(qPCR_means)
```

plotting 
```{r, message=FALSE, warning=FALSE,tidy=TRUE}
Fig1B <- ggplot2::ggplot() +
  geom_bar(data=qPCR_means, aes(x=Target, y=log2(Mean_Fold_Change+1), fill=Target), stat="identity", color="black", alpha=0.9) +
  geom_point(data=qPCR, aes(x=Target, y=log2(Fold_Change+1), fill=Target), size=3.5, pch=21, color="black") +
  facet_wrap(~Substrate) + 
  theme_pubr() + 
  theme(axis.text.x = element_text(angle=90),
        legend.position = "none") +
  ylim(0,7) +
  scale_fill_manual(values=c("#C7E8BF", "#8FCF91" , "#197B41"))

#pdf(file = "/Users/riccardomosca/Desktop/RAPseq_PAPER/FIGUREs/FIGURE1/NEW/HUR_RAPqPCR_barplot_Log2.pdf",2.5,5)
#dev.off()
Fig1B 




```

Figure S1E
Loading packages and annotated peak files, keeping the unique targets
```{r, message=FALSE, warning=FALSE,tidy=TRUE}
library(eulerr)
library(tidyverse)
library(clusterProfiler)
library(org.Hs.eg.db)
library(reshape2)
library(ggvenn)
library(dplyr)


path <- "/Users/riccardomosca/Desktop/RAPseq_PAPER/FIGUREs/FIGURE1/NEW/HUR_hnRNPA1_titration_revision_231016/DEEPER_SEQ/PEAKs"

# Different HUR titrations 
types <- c("HUR100", "HUR50", "HUR25")

for (t in types) {
  # build the filename and read it in
  infile <- file.path(path, paste0(t, "_annotated.txt"))
  peaks  <- read.delim(infile, sep = "\t", header = TRUE)
  
  # unique RNAs
  df <- peaks %>%
    dplyr::select(gene_name, Gene_BS, gene_ID, positive_fa, negative_fa) %>%
    filter(!duplicated(gene_name)) %>%
    as.data.frame()
  
  # rename first column to "gene"
  colnames(df)[1] <- "gene"
  
  
  assign(paste0("", t), df)
}

#Reading HUR eCLIP bed file from ENCODE https://www.encodeproject.org/experiments/ENCSR090LNQ/
eCLIP_HUR_peaks <- read.delim("/Users/riccardomosca/Desktop/RAPseq_PAPER/FIGUREs/FIGURE1/NEW/HUR_hnRNPA1_titration_revision_231016/PEAKs/HUR_eCLIP_annotated_peaks.txt", sep = '\t', header = T)
HUR_eCLIP <- eCLIP_HUR_peaks %>% dplyr::select(gene_name) 
HUR_eCLIP <- HUR_eCLIP[!duplicated(HUR_eCLIP$gene_name),]
HUR_eCLIP <- as.data.frame(HUR_eCLIP) 
colnames(HUR_eCLIP)[1] <- "gene"


```


```{r,message=FALSE, warning=FALSE,tidy=TRUE}
#Venn diagram
gene_HUR100 <- HUR100$gene 
gene_HUR50 <- HUR50$gene 
gene_HUR25 <- HUR25$gene 
gene_HUR_eCLIP <- HUR_eCLIP$gene

ABCD <- list(HUR100 = gene_HUR100, HUR50 = gene_HUR50, HUR25 = gene_HUR25, HUR_eCLIP = gene_HUR_eCLIP)

ggvenn(
  ABCD, 
  fill_color = c("#09316A", "#4292C6", "#9ECAE0", "#187D41"),
  stroke_size = 0.5, set_name_size = 4
)
```

Computing the mean of Gene BS for the three titrations overlap, and each specific one. 
```{r, message=FALSE, warning=FALSE,tidy=TRUE}
overlap_HUR100_HUR50 <- subset(HUR100, gene %in% HUR50$gene) 
overlap_HUR50_HUR25 <- subset(HUR50, gene %in% HUR25$gene) 
overlap_HUR100_HUR50_HUR25 <- subset(overlap_HUR100_HUR50, gene %in% overlap_HUR50_HUR25$gene)
specific_HUR100 <-HUR100[!as.character(HUR100$gene) %in% as.character(HUR50$gene),]
specific_HUR100 <-specific_HUR100[!as.character(specific_HUR100$gene) %in% as.character(HUR25$gene),]
specific_HUR50 <-HUR50[!as.character(HUR50$gene) %in% as.character(HUR100$gene),]
specific_HUR50 <-specific_HUR50[!as.character(specific_HUR50$gene) %in% as.character(HUR25$gene),] 
specific_HUR25 <-HUR25[!as.character(HUR25$gene) %in% as.character(HUR100$gene),]
specific_HUR25 <-specific_HUR25[!as.character(specific_HUR25$gene) %in% as.character(HUR50$gene),]


# Determine the maximum length of the vectors to combine
max_length <- max(length(overlap_HUR100_HUR50_HUR25$Gene_BS), length(specific_HUR100$Gene_BS), length(specific_HUR50$Gene_BS), length(specific_HUR25$Gene_BS))

# Create vectors of equal length by recycling shorter vectors
Gene_BS_HUR100_50_25 <- c(overlap_HUR100_HUR50_HUR25$Gene_BS, rep(NA, max_length - length(overlap_HUR100_HUR50_HUR25$Gene_BS)))
Gene_BS_HUR100 <- c(specific_HUR100$Gene_BS, rep(NA, max_length - length(specific_HUR100$Gene_BS)))
Gene_BS_HUR50 <- c(specific_HUR50$Gene_BS, rep(NA, max_length - length(specific_HUR50$Gene_BS)))
Gene_BS_HUR25 <- c(specific_HUR25$Gene_BS, rep(NA, max_length - length(specific_HUR25$Gene_BS)))

# Making an unique df
combined_BS <- data.frame(BS_overlap = Gene_BS_HUR100_50_25, BS_HUR100 = Gene_BS_HUR100, BS_HUR50 = Gene_BS_HUR50, BS_HUR25 = Gene_BS_HUR25)
combined_BS <- melt(combined_BS, value.name = "BS", variable.name = "Group")
combined_BS <- na.omit(combined_BS)

head(combined_BS)

combined_BS %>% 
  group_by(Group) %>%
  summarise(n=n())

aa <- ggplot( data=combined_BS, aes(x=Group, y=log2(BS+1)) ) +
  geom_jitter(aes(color=Group), pch=16, alpha=0.25) +
  scale_colour_manual(values = c("#000000","#0A316A", "#4292C6", "#9ECAE0" )) +
  theme_classic(base_size = 12.5) +
  theme(legend.position="none")

FigS1E <- aa + geom_violin(trim = T, bw=0.75, scale="width", lwd=1, fill=NA) +
  geom_boxplot(outlier.shape = NA, width=0.15, lwd=1, fill=NA) 

FigS1E



# collect specific Gene BS vectors into a named list
specific_list <- list(
  HUR100 = specific_HUR100$Gene_BS,
  HUR50  = specific_HUR50$Gene_BS,
  HUR25  = specific_HUR25$Gene_BS
)

# run wilcox.test() for each, comparing to the overlap vector
wt <- lapply(names(specific_list), function(nm) {
  y   <- specific_list[[nm]]
  test <- wilcox.test(overlap_HUR100_HUR50_HUR25$Gene_BS, y)
  data.frame(
    comparison = paste0("overlap_vs_", nm),
    W          = as.numeric(test$statistic),
    p.value    = test$p.value
  )
})

# combine into one df
wt_df <- do.call(rbind, wt)




```

